### 2018-07-09 methods.tex
Reprocessing logic and table added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h!]
        %\centering
        \caption{Summary of principal data for MSBR \cite{robertson_conceptual_1971}.}
        \begin{tabularx}{\textwidth}{ s  s}
        \hline
                Thermal capacity of reactor           & 2250 MW(t)
                \\ 
                Net electrical output                 & 1000 MW(e) 
                \\  
                Net thermal efficiency        & 44.4\%
                \\  
                Salt volume fraction in central core zone     & 0.13
                \\ 
                Salt volume fraction in outer core zone       & 0.37
                \\ 
                Fuel salt inventory (Zone I)                  & 8.2 m$^3$	
                \\ 
                Fuel salt inventory (Zone II)                 & 10.8 m$^3$	
                \\ 
                Fuel salt inventory (annulus)                 & 3.8 m$^3$	
                \\  
                Total fuel salt inventory                     & 48.7 m$^3$	
                \\ 
                Fissile mass in fuel salt                   & 1303.7 kg	
                \\ 
                Fuel salt components                  & 
                LiF-BeF$_2$-ThF$_4$-$^{233}$UF$_4$	
                \\  
                Fuel salt composition                 & 
                71.75-16-12-0.25 mole\%
                \\
                Fuel salt density                    & 
                3.35 g/cm$^3$
                \\ \hline
        \end{tabularx}
        \label{tab:msbr_tab}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Online reprocessing method}
Removing specific chemical elements from a molten salt is a complicated task that requires intelligent design (e.g., chemical separations equipment design, fuel salt flows to equipment) and has a considerable economic cost.

\subsection{The SaltProc modeling and simulation code}

The SaltProc tool \cite{andrei_rykhlevskii_arfc/saltproc:_201*} is designed to expand SERPENT 2
 depletion capabilities for modeling liquid-fueled \gls{MSR} for continuous reprocessing.
The Python package uses HDF5 \cite{the_hdf_group_hierarchial_1997} to store data, and the Nuclear Engineering Toolkit PyNE \cite{scopatz_pyne:_2012}
for SEPRENT output file parsing and nuclide naming. It is an open-source tool that uses a semi-continuous approach to 
simulate continuous feeds and removals in the \gls{MSR}.

The tool structure and capabilities of SaltProc are similar to ChemTriton tool for SCALE
developed in \gls{ORNL} \cite{powers_new_2013}. SaltProc is coupled with Monte Carlo SERPENT 2
software which allows to simulate online reprocessing for irregular full-core geometry with 
high level of fidelity.  The primary function of SaltProc is to manage material streams while
SERPENT 2 performs most of the computationally heavy work, namely neutron transport and depletion
calculations. Saltproc is defined as a 
python class, where each material stream is defined as a isotopic atomic density
vector variable. This allows tracking of time-sensitive material streams such as the
Pa-233 tank in the \gls{MSBR}. The user can define the reprocessing parameters, such as the 
reprocessing interval and removal efficiency.  In addition, SaltProc provides a set of functions 
for each stream: read and write isotopic data in/from database, separate out specific isotopes 
from stream with defined efficiency, feed in specific isotopes to stream, and maintain constant 
number density of specific nuclide in the core. These attributes and functions are crucial to 
simulate the operation of a complex, multi-zone, multi-fluid \gls{MSR} and are universal for 
myriad reactor systems.

Saltproc is currently in active development in Github (https://github.com/arfc/saltproc), and has unit tests and
Travis continuous integration for sustainable development. There is also documentation
generated through Sphinx document generator for ease of use. We plan to implement
user-defined reprocessing scheme definition, two-region \gls{MSR} modeling capabilities,
and decay modeling in future releases.

Figure~\ref{fig:saltproc_flow} illustrates the  online reprocessing simulation algorithm the 
coupled SaltProc and SERPENT 2. To perform a depletion step,
SaltProc reads an external SERPENT 2 template file which is defined by the user. This file 
contains input cards with parameters such as geometry,
moderator and construction materials isotopic composition, neutron population, criticality 
cycles, total heating power, and boundary conditions.
After the depletion calculation, SaltProc reads the depleted fuel composition file (\texttt{
.bumat1}) and stores the depleted
composition isotopic vector in an HDF5 database. SaltProc only stores and edits the isotopic 
composition of the fuel stream,
which makes SaltProc a flexible tool to model any geometry: an infinite medium, a unit cell, a 
multi-zone simplified assembly, or a full core.
This flexibiliity allows the user to perform simulations of varying fidelity and computational 
intensity.

SaltProc can manage as many material streams as desired. It also may work with multiple depletion 
materials. At the end of a each depletion
step, SaltProc reads the depleted compositions and tracks each material stream individually. 
Following this, it applies chemical
separation functions to fuel stream vectors. These vectors then form a matrix (isotopics x 
timesteps) which SaltProc stores in an HDF5 database
and prints into the SERPENT 2 composition file for the next depletion calculation.

Saltproc datasets are timseries, meaning that every value is recorded every timestep. The 
datasets SaltProc produces are listed below,
where the values inside the parenthesis are the dataset sizes:

\begin{itemize}
    \item \texttt{core adensity before reproc} (number of isotopes X timesteps)
    \item \texttt{core adensity after reproc} (number of isotopes X timesteps)
    \item \texttt{Keff_BOC} (1 X timesteps)
    \item \texttt{Keff_EOC} (1 X timesteps)
    \item \texttt{Th tank adensity} (number of isotopes X timesteps)
    \item \texttt{iso codes} (number of isotopes X 1)
\end{itemize}

Liquid-fueled \gls{MSBR} design in this thesis focuses on the state of the core at an equilibrium condition, after fission products have
built up in the fuel salt during years if operation. Isotopic composition of the fuel salt continues change slightly even after decades
of operation, but the dominant nuclides that have significant impact to the neutronic behavior tend to reach an equilibrium concentration
(e.g., vary less than 1\% over several years). In contrast, from the startup of an \gls{MSBR} until the equilibrium condition,
the fuel salt composition undergoes significant changes (e.g., changes in fission products, minor actinides, and fissile
materials number density). During this period, the material feeds and removals should be optimized for the fastest \gls{MSBR} transition to an
equilibrium state. A faster transition simplifies the reactor operation because, at equilibrium, the fissile and fertile feed rates,
fission product removal rates, and corresponding core safety parameters are constant in time.

\begin{figure}[htp!] % replace 't' with 'b' to 
  \centering
  \vspace{-0.3em}
  \includegraphics[width=\textwidth]{saltproc_flowchart.pdf}
  \caption{Flow chart for the Saltproc python-based tools.}
  \vspace{-1.0em}
  \label{fig:saltproc_flow}
\end{figure}
\FloatBarrier

In addition, SaltProc is able to define time-dependent material feed and removal rates to investigate the their impacts. These rates need not be
constant in SaltProc. They can be defined as piecewise functions or set to respond conditions in the core. For instance, SaltProc might increase the
fissile material feeding rate if the effective multiplication factor, $k_{eff}$, falls below a specific limit (e.g., 1.002).
These capabilities allow SaltProc to analyze fuel cycle of a generic liquid-fueled \gls{MSR}. In summary, the development approach of SaltProc focused
on producing a generic, flexible and expandable tool to give the SERPENT 2 Monte Carlo code the ability to conduct advanced in-reactor
fuel cycle analysis as well as simulate a myriad of online refueling and fuel reprocessing systems.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h!]
        %\centering
        \caption{Summary of principal data for MSBR \cite{robertson_conceptual_1971}.}
        \begin{tabularx}{\textwidth}{ s  s}
        \hline
                Thermal capacity of reactor           		& 2250 MW(t)
                \\ 
                Net electrical output                 		& 1000 MW(e) 
                \\  
                Net thermal efficiency        				& 44.4\%
                \\  
                Salt volume fraction in central zone I		& 0.13
                \\ 
                Salt volume fraction in outer zone II       & 0.37
                \\ 
                Fuel salt inventory (Zone I)                & 8.2 m$^3$	
                \\ 
                Fuel salt inventory (Zone II)               & 10.8 m$^3$	
                \\ 
                Fuel salt inventory (annulus)               & 3.8 m$^3$	
                \\  
                Total fuel salt inventory                   & 48.7 m$^3$	
                \\ 
                Fissile mass in fuel salt                   & 1303.7 kg	
                \\ 
                Fuel salt components                  & 
                LiF-BeF$_2$-ThF$_4$-$^{233}$UF$_4$	
                \\  
                Fuel salt composition                 & 
                71.75-16-12-0.25 mole\%
                \\
                Fuel salt density                    & 
                3.35 g/cm$^3$
                \\ \hline
        \end{tabularx}
        \label{tab:msbr_tab}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Online reprocessing method}
Removing specific chemical elements from a molten salt is a complicated task that requires intelligent design (e.g., chemical separations equipment design, fuel salt flows to equipment) and has a considerable economic cost. All liquid-fueled \gls{MSR} designs involve varying levels of online fuel processing. Minimally, volatile gaseous fission products (e.g. Kr, Xe) escape from the fuel salt during routine reactor operation and must be captured. Additional systems might be used to enhance removal of those elements. Most designs also call for the removal of noble and rare earth metals from the core since these metals act as neutron poisons. Some designs suggest a more complex list of elements to process (figure ~\ref{fig:periodic_tab}), including the temporary removal of protactinium from the salt or other regulation of the actinide inventory in the fuel salt \cite{ahmad_neutronics_2015}.

In the single-fluid \gls{MSBR} considered in this work, thorium, uranium, protactinium, and fission products are all mixed together in a single fluoride salt (FLiBe). Separation of thorium from lanthanide (atomic numbers 57 through 71) fission products is rather challenging because of their chemical similarities.

\begin{figure}[htp!] % replace 't' with 'b' to 
  \centering
  \includegraphics[width=\textwidth]{periodic_map.png}
  \caption{Processing options for \gls{MSR} fuels \cite{ahmad_neutronics_2015}.}
  \label{fig:periodic_tab}
\end{figure}

\subsubsection{Fuel material flows}
The $^{232}$Th in the fuel absorbs thermal neutrons and produces $^{233}$Pa which then decays into the fissile $^{233}$U. Furthermore, the \gls{MSBR} design requires online reprocessing to remove all poisons (e.g. $^{135}$Xe), noble metals, and gases (e.g. $^{75}$Se, $^{85}$Kr) every 20 seconds. Protactinium presents a challenge, since it has a large absorption cross section in the thermal energy spectrum. Accordingly, $^{233}$Pa is continuously removed from the fuel salt into a protactinium decay tank to allow $^{233}$Pa to decay to $^{233}$U without poisoning the reactor. The reactor reprocessing system is designed to separate $^{233}$Pa from the molten-salt fuel over 3 days, hold it while $^{233}$Pa decays into $^{233}$U, and return it back to the primary loop. This feature allows the reactor to avoid neutron losses to protactinium, keeps fission products to a very low level, and increases the efficiency of $^{233}$U breeding. Table~\ref{tab:reprocessing_list} summarizes full list of nuclides and the cycle times used for modeling salt treatment and separations \cite{robertson_conceptual_1971}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[ht!]
        \centering
        \caption{The effective cycle times for protactinium and fission products removal (reproduced from \cite{robertson_conceptual_1971}).}
        \begin{tabularx}{\textwidth}{ x | s | x }
        \hline 
        Processing group & \qquad\qquad\qquad Nuclides & Cycle time (at full power) \\ \hline 
        Rare earths & Y, La, Ce, Pr, Nd, Pm, Sm, Gd & 50 days \\ 
        \qquad & Eu & 500 days \\ 
        Noble metals & Se, Nb, Mo, Tc, Ru, Rh, Pd, Ag, Sb, Te & 20 sec \\
        Seminoble metals & Zr, Cd, In, Sn & 200 days \\
        Gases & Kr, Xe & 20 sec \\ 
        Volatile fluorides & Br, I & 60 days \\
        Discard & Rb, Sr, Cs, Ba & 3435 days \\ 
        Salt discard & Th, Li, Be, F & 3435 days \\ 
        Protactinium & $^{233}$Pa & 3 days \\ 
        Higher nuclides & $^{237}$Np, $^{242}$Pu & 16 years \\ 
        \end{tabularx}
        %\footnotetext{Chemical processing plant and gas separation system removing chemical elements (not isotopes) only. Isotopes instead of elements listed because other isotopes are short-lived and might be ignored.}
        \label{tab:reprocessing_list}
\end{table}
Th removal rates vary among nuclides in this reactor concept which dictate the necessary resolution of depletion calculations. If the depletion time intervals are very short, an enormous number of depletion steps are required to obtain the equilibrium composition. On the other hand, if the depletion  calculation time interval is too long, the impact of short-lived fission products is not captured. To compromise, the time interval for depletion calculations in this model was selected as 3 days to correlate with the removal interval of $^{233}$Pa and $^{232}$Th was continuously added to maintain the initial mass fraction of $^{232}$Th.

\subsubsection{The SaltProc modeling and simulation code}
